---
title: "week3 regression"
author: "Astha Agarwal"
date: "11/11/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Report Summary

The aim of this reoprt is to analyze the relation bewteen the transmission type of the vehicle(automatic or manual)and the distance(in miles) as we all know it covers per gallon(mpg).
The final result aims at determe as to which transmission type gives more miles per gallon rating is v low.
The `mtcars` dataset as we are using is used to make the analysis.

#### Load Data
Load the dataset and convert categorical variables to factors.
```{r results='hide', message=FALSE}
library(ggplot2)
data(mtcars)
head(mtcars, n=3)
dim(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- factor(mtcars$am)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
attach(mtcars)
```

#### Exploratory Analysis
**See Appendix Figure I**The graph leads us to believe that there is a significant 
increase in MPG when for vehicles with a manual transmission vs automatic.

##### Statistical Inference
T-Test transmission type and MPG
```{r}
trs <- t.test(mpg ~ am)
trs$p.value
```
The T-Test rejects the null hypothesis that we have come accross that the difference between transmission
types is 0.  
```{r}
trs$estimate
```
The difference estimate between the given 2 transmissions is 7.2449 MPG in favor of the 
manual.

##### Regression Analysis
Fit the full model of the data
```{r results='hide'}
fmft <- lm(mpg ~ ., data = mtcars)
summary(fmft)  # results hidden
summary(fmft)$coeff  # results hidden
```
Since none of the coefficients that we have found have a p-value less than 0.05 we cannot conclude that we have found out
which variables are more statistically significant. 

Backward selection to determine whi other variables are most statistically 
significant than any othr
```{r results='hide'}
sft <- step(fmft)
summary(sft) # results hidden
summary(sft)$coeff # results hidden
```

The new model has 4 variables (cylinders, horsepower, weight, transmission). The
R-squared value of 0.8659 confirms that this model explains about 87% of the 
variance in MPG. The p-values also as we have known are statistically significantly because they
have a p-value less than 0.05. The coefficients conclude from the gven above that increasing the 
number of cylinders from 4 to 6 with decrease the the the MPG by 3.03.  Further 
increasing the cylinders to 8 with the utmost decrease the MPG by 2.16.  Increasing the 
horsepower is decreases MPG 3.21 for every 100 horsepower.  Weight decreases the
MPG by 2.5 for each 1000 lbs increase. A Manual transmission improves the MPG by
1.81.

#### Residuals & Diagnostics
Residual Plot
**See Appendix Figure II**

The plots conclude:

A: The randomness supports the assumption that we have made of independence  
B: The distribution of all residuals is normal  
C: Confirms the constant of the variance assumption  
D: Since all the points that we have seen are under 0.05, there are no outliers
```{r}
sum((abs(dfbetas(sft)))>1)
```

#### Results

The analysis performed shows that the vehicles with manual transmission have a higher MPG rating than vehicles with automatic transmission by 7.245.
When fitted with multiple linear regressions, the results went on to show that the transmission type had less to do with the earlier results, that is, the manual transmission itself contributed less to MPG, still giving it an edge over automatic transmission, but by 1.81 MPG. Other variables like it's enigne capacity(horsepower), it's weight and number of cylinders had a greater effect on the MPG of vehicles.  

#### Conclusion

There is a difference given in the observed  MPG based on transmission type.  

#### Appendix Figures

#### I
Exploratory Box graph has een shwn fr whatthat compares Automatic and Manual 
transmission MPG.  
```{r echo=FALSE}
  boxplot(mpg ~ am, 
          xlab="Type if Transmission (0 = Automatic, 1 = Manual)", 
          ylab="mpg",
          main="MPG by  Vehicle Transmission Type")
```
part 2 as shown heree
#### II 
```{r echo=FALSE}
par(mfrow = c(2, 2))
plot(sft)
```